Final Model Analysis

### 2017-08-27 model fitting/final_model.Rmd
started writing function to predict survival probabilities

* new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strata(ampm) + n_images + sqrt(avg_tag_score) + poly(text_length, degree = 2) + pspline(device_length, df = 0, method = 'AIC') + pspline(avg_tag_length, df = 0, method = 'AIC') + pspline(newline_ratio, df = 0, method = 'AIC')

@@

```{r, echo = FALSE}
# function to perform cross validation, vars should be string: "new_category + new_user..."
crossval <- function(vars, train, test) {
  formula <- paste("Surv(time_until_answer, answered) ~ ", vars, sep = "")
  model <- rms::cph(as.formula(formula), data = train)

@@

get_avgmetrics(cv_results)

get_avgdiff(cv_results) # metrics don't change that much between fitting model on training and test data sets
```

```{r, echo = FALSE}
# validation and calibration with rms package

valmodel <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = x, x = TRUE, y = TRUE, surv = TRUE, time.inc = 0.5, dxy = TRUE)

validation <- rms::validate(calmodel, method = "crossvalidation", B = 100, bw = F, dxy = TRUE, u = 0.5)
validation
# output: 
# Dxy = Somers' Dxy (2 ∗ (C − 0.5), where C is the concordance probability)
# R2: RN2 index
# Slope: Calibration slope (slope of predicted log odds vs true log odds) optimism value is a value of overfitting
# D: Discrimination index — likelihood ratio χ2 divided by the sample size
# U: Unreliability index — unitless index of how far the logit calibration curve intercept and slope are from (0, 1)
# Q: Logarithmic accuracy score - scaled version of the log-likelihood achieved by the predictive model
# g: g index (?)

# validate the model for calibration accuracy in predicting probability at 0.5 hours 
calibration <- rms::calibrate(calmodel, method = "crossvalidation", B = 100, bw = FALSE, u = 0.5)
calibration
plot(calibration)
# predicts and calibrates at 0.5 hours for each iteration of the bootstrap
# plot gives mean absolute error in predictions, mean squared error, and the 0.9 quantile of the absolute error
# error =  difference between the predicted values and the corresponding bias-corrected calibrated values

@@

@@

```{r, echo = FALSE}
# getting estimated survival probabilites
surv_prob <- model$surv # list of 4 (one list for each stratification level)

@@

# gives one survival probability (not split up into strat levels)
# either specify newdata, linear.predictors or x
# number on the rows of the surv matrix is the index 
# function is very slow if full data set is used
rms::survest(model, newdata = test1[1:3,], times = c(0.5, 1, 10), conf.int = FALSE)

# using different function
pec::predictSurvProb(model, newdata = test1[1:3,], times = c(0.5, 1, 10))
```

>>>>

* new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strata(ampm) + sqrt(avg_tag_score) + poly(text_length, degree = 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4)
* device_length has 5 knots since hypothesized that it's a better predictor, others only have 4 (reccomended in Regression Modelling Strategies text)

@@

```{r, functions for cv, echo = FALSE}
# function to perform cross validation, vars should be string: "new_category + new_user..."
crossval <- function(vars, train, test) {
  formula <- paste("Surv(time_until_answer, answered) ~ ", vars, sep = "")
  model <- rms::cph(as.formula(formula), data = train)

@@

get_avgmetrics(cv_results)
get_avgdiff(cv_results) # metrics don't change that much between fitting model on training and test data sets
```

##### Validation with rms::validate for Somers’ Dxy rank correlation between pre-dicted log hazard and observed survival time, + slope shrinkage

###### output: 
* Dxy = Somers' Dxy (2 ∗ (C − 0.5), where C is the concordance probability)
* R2: RN2 index
* Slope: Calibration slope (slope of predicted log odds vs true log odds) optimism value is a value of overfitting
* D: Discrimination index — likelihood ratio χ2 divided by the sample size
* U: Unreliability index — unitless index of how far the logit calibration curve intercept and slope are from (0, 1)
* Q: Logarithmic accuracy score - scaled version of the log-likelihood achieved by the predictive model
* g: g index (?)

###### interpreting:
* optimism is small- model is a good fit (?)
* slope shrinkage = 0.0163, indicates over-fitting not present

```{r, echo = FALSE}
valmodel <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = x, x = TRUE, y = TRUE, surv = TRUE, time.inc = 0.5, dxy = TRUE)

validation <- rms::validate(valmodel, method = "crossvalidation", B = 100, bw = F, dxy = TRUE, u = 0.5)
validation

@@

##### Validating the model for calibration accuracy with rms::calibrate

* outputs mean absolute error in predictions, mean squared error, and the 0.9 quantile of the absolute error
* error =  difference between the predicted values and the corresponding bias-corrected calibrated values

```{r, echo = FALSE}
# calibration accuracy in predicting probability at 0.5 hours (predicts and calibrates at 0.5 hours for each iteration of the bootstrap?)
calibration <- rms::calibrate(calmodel, method = "crossvalidation", B = 100, bw = FALSE, u = 0.5)
calibration
plot(calibration)
```

@@

```{r, echo = FALSE}
surv_prob <- model$surv # list of 4 (one list for each stratification level)

@@

# either specify newdata, linear.predictors or x
# number on the rows of the surv matrix is the index 
# function is very slow if full data set is used
rms::survest(model, newdata = test1[1:3,], times = c(0.5, 1, 10), conf.int = FALSE)

# using different function
pec::predictSurvProb(model, newdata = test1[1:3,], times = c(0.5, 1, 10))

times <- sort(unique(x$time_until_answer))

pec::plotPredictSurvProb(model, newdata = x[1:4,], times = times) # this is ugly, figure out how to do this in ggplot?
```

##### Describing the final model

* predicted survival curve + curves for certain variables (rms functions to do this keep outputting errors about datadist)
* nomograms?

##### Function to output survival probabilities

* input data set with any combination of predictors + times to predict on 
* output corresponding survival probabilities

```{r}
# input vector of times to predict on + data set to be used in predictions
# function runs really slowly because it has to set up the data + fit the model everytime 
# outputs matrix of predicted survival 
# also include some kind of performance metric? measure of how accurately its predicting 
# sets up variables needed in the new data 
get_survival <- function(times, data) {
  newdata <- oshitar::variable_setup(data)
  x <- oshitar::setup()
  model <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = x, x = TRUE, y = TRUE, surv = TRUE)
  surv_probs <- pec::predictSurvProb(model, newdata = newdata, times)
  return(surv_probs)
}

get_survival(times = c(0.5, 1, 10), data = out)
